---
title: "The 'RMS Titanic' Survivors"
author: "Tomás A. Maccor"
date: "15/4/2020"
output: 
    slidy_presentation:
    theme: readable
---


## App Overview

This Shiny app enables the user to analyze fate of the **RMS Titanic's** passengers & crew.
via exploratory plotting of the passenger survival data.   
  
The data used by the App comes from one of R's built in datasets  --Titanic.\  
But it can very easily be modified to accept other dataset & plot them.\  
\    
It downloads the dataset & converts it to a dataframe so that it can be later passed on to the FILTER command from the DPLYR package


```{r, echo=TRUE}
library(shiny)
library(shinythemes)
library(dplyr, quietly = TRUE , warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE)
library(readr, quietly = TRUE)

data("Titanic")
Titanic <- as.data.frame(Titanic)

```

The dataset has 5 variables and 32 observations:

```{r, echo=FALSE}
a <- Titanic
knitr::kable(head(a), align = c("l") )
```

\break


## App Layout


* The App has a SIDEBAR Panel that has 2 sliders where the user will input the visualization choices
* The user can choose the visualise all people that were onboard of the ship, OR by passenger category
    + 1st class passengers
    + 2nd class passengers
    + 3rd class passengers
    + ship's crew
* PLUS the user can choose if they want to visualize either gender at one given time, or both
\  
* The MAIN Panel displays the plot of the user selection
* It also displays the sum of the total # of passengers contained in the user selection

\break

## R code and Shiny commands used for the App

The UI code accepts User Input via 2 "selectInput" Shiny commands:   
* PassengerClass    
* Sex

```{r, echo=TRUE, eval=FALSE}
selectInput(inputId = "PassengerClass", label = "Enter passenger category",
                        choices = c(levels(Titanic$Class), "All")  ),
                        
            selectInput("Sex", "Choose sex", choices = c(levels(Titanic$Sex), "Both")   ),
            submitButton(text = "Show Plot")
```

The Shiny SELECT boxes were improved in their functionality, so that they can also accept an input of ALL categories of the PASSENGER CLASS & SEX variables from the dataset.
\  
\  

The algorithm in the **server.R** file used to subset the Titanic dataset (based on the user selection) is provided below:

```{r, echo=TRUE, eval=FALSE}

ifelse (input$PassengerClass =="All", Titanic_subset <- Titanic, 
                     Titanic_subset <- Titanic %>% filter(Class == input$PassengerClass)
                     )
              
              if (input$Sex == "Male") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Male")}
              if (input$Sex == "Female") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Female")}
              if (input$Sex == "Both") {Titanic_subset <- Titanic_subset}
              
              ploteo <- ggplot(Titanic_subset, aes(x = Class, y = Freq, fill = Survived ) ) + 
                               geom_bar(stat = "identity", position = position_dodge() ) +
                               ylab("Number of passengers")
                              
              ploteo


```


## Exploratory analysis

By making exploratory plots, we can quickly obtain insights from this dataset.     
    
For example, the overall survival for 2nd Class passenger tells us that 1/3 of the passengers survived, but if we differentiate by sex we will see that the % of females that survived was roughly inversely proportional to the male survivors. Thus, we can confirm that the "women first" call was indeed observed for the 2nd Class passengers.      

**ALL 2nd Class passengers**  
*User input PASSENGERCLASS = "All"*   
*User input SEX = "Both"*

```{r, echo=FALSE, message=FALSE}

inputPassenger <- "2nd"
inputSex <- "All"

ifelse (inputPassenger =="All", Titanic_subset <- Titanic, 
       Titanic_subset <- filter(Titanic, Class == inputPassenger)
       )
       
        if (inputSex == "Male") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Male")}
        if (inputSex == "Female") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Female")}
        if (inputSex == "Both") {Titanic_subset <- Titanic_subset}
        
        ploteo <- ggplot(Titanic_subset, aes(x = Class, y = Freq, fill = Survived ) ) + 
            geom_bar(stat = "identity", position = position_dodge() ) +
            ylab("Number of passengers")
        
        ploteo
```

**2nd Class FEMALE passengers**

```{r, echo=FALSE, message=FALSE}
inputPassenger <- "2nd"
inputSex <- "Female"

ifelse (inputPassenger =="All", Titanic_subset <- Titanic, 
       Titanic_subset <- filter(Titanic, Class == inputPassenger)
       )
       
        if (inputSex == "Male") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Male")}
        if (inputSex == "Female") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Female")}
        if (inputSex == "Both") {Titanic_subset <- Titanic_subset}
        
        ploteo <- ggplot(Titanic_subset, aes(x = Class, y = Freq, fill = Survived ) ) + 
            geom_bar(stat = "identity", position = position_dodge() ) +
            ylab("Number of passengers")
        
        ploteo



```


**2nd Class MALE passengers**

```{r, echo=FALSE}
inputPassenger <- "2nd"
inputSex <- "Male"

ifelse (inputPassenger =="All", Titanic_subset <- Titanic, 
       Titanic_subset <- filter(Titanic, Class == inputPassenger)
       )
       
        if (inputSex == "Male") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Male")}
        if (inputSex == "Female") {Titanic_subset <- Titanic_subset %>% filter(Sex == "Female")}
        if (inputSex == "Both") {Titanic_subset <- Titanic_subset}
        
        ploteo <- ggplot(Titanic_subset, aes(x = Class, y = Freq, fill = Survived ) ) + 
            geom_bar(stat = "identity", position = position_dodge() ) +
            ylab("Number of passengers")
        
        ploteo



```
